---
title: "Antidiabetic Nontreatment Timelines: `r params$target_drug` vs. `r params$nontreatment_group`"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    df_print: kable
params:
  target_drug:        "Semaglutide"
  nontreatment_group: "Nontreatment"
  result_file:        "Data/nontreatment_timelines_result-Semaglutide.rds"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(scales)

load(params$result_file)
r <- nontreat_result
```

# Eligible Population Summary

```{r population-summary}
cat(sprintf(
    "There are %s individuals eligible for the %s population and %s in the %s population. Match ratio of %d to 1.",
    comma(r$n_comp),
    params$nontreatment_group,
    comma(r$n_sema),
    r$target_drug,
    r$match_ratio
))
```

---

# Treatment Group: Time from Eligibility to Index

Distribution of days from MDD diagnosis to index date in the **`r params$target_drug`** group.

```{r tfe-dist-plot, fig.width=12, fig.height=5}
ggplot(r$tfe_dist, aes(x = tfe_at_index_days, y = n)) +
    geom_bar(stat = "identity", fill = "salmon") +
    theme_minimal(base_size = 14) +
    labs(
        title = paste0(r$target_drug, ": Time from MDD Diagnosis to Index Date"),
        x     = "Days from MDD Diagnosis to Index",
        y     = "Count"
    )
```

---

## TFE Distribution: Treatment vs. Comparison

```{r tfe-comparison-plot, fig.width=12, fig.height=5}
ggplot(r$dist_comp, aes(x = tfe_at_index_days, y = n, fill = group)) +
    geom_bar(stat = "identity") +
    scale_fill_viridis_d(option = "plasma", end = 0.7) +
    theme_minimal(base_size = 14) +
    labs(
        title = "Time from MDD Diagnosis to Index Date",
        x     = "Days from MDD Diagnosis to Index",
        y     = "Count",
        fill  = "Group"
    )
```

## KS Test: TFE Distribution

```{r ks-tfe}
print(r$ks_tfe)
```

---

# Final Cohort Diagnostics

## Age at Index

```{r age-plot, fig.width=12, fig.height=5}
ggplot(r$dte_cohort_data3, aes(x = age_at_index_years)) +
    geom_histogram(aes(y = after_stat(density),
                       color = treatment_name,
                       fill  = treatment_name),
                   alpha    = 0.4,
                   position = "dodge") +
    geom_density(aes(color = treatment_name)) +
    scale_fill_viridis_d(option = "plasma", end = 0.7) +
    scale_color_viridis_d(option = "plasma", end = 0.7) +
    theme_minimal(base_size = 14) +
    labs(
        title = "Age at Index Date",
        x     = "Age (years)",
        y     = "Density",
        fill  = "Group",
        color = "Group"
    )
```

**KS test — age at index:**

```{r ks-age}
print(r$ks_age)
```

## Time from Diagnosis to Index

```{r tdi-plot, fig.width=12, fig.height=5}
ggplot(r$dte_cohort_data3, aes(x = time_diag_to_index_days)) +
    geom_histogram(aes(y = after_stat(density),
                       color = treatment_name,
                       fill  = treatment_name),
                   alpha    = 0.4,
                   position = "dodge") +
    geom_density(aes(color = treatment_name)) +
    scale_fill_viridis_d(option = "plasma", end = 0.7) +
    scale_color_viridis_d(option = "plasma", end = 0.7) +
    theme_minimal(base_size = 14) +
    labs(
        title = "Time from MDD Diagnosis to Index Date",
        x     = "Days",
        y     = "Density",
        fill  = "Group",
        color = "Group"
    )
```

**KS test — time from diagnosis to index:**

```{r ks-tdi}
print(r$ks_tdi)
```

## Index Year

```{r year-plot, fig.width=12, fig.height=5}
ggplot(r$dte_cohort_data3, aes(x = index_year)) +
    geom_histogram(aes(y = after_stat(density),
                       color = treatment_name,
                       fill  = treatment_name),
                   alpha    = 0.4,
                   position = "dodge") +
    geom_density(aes(color = treatment_name)) +
    scale_fill_viridis_d(option = "plasma", end = 0.7) +
    scale_color_viridis_d(option = "plasma", end = 0.7) +
    theme_minimal(base_size = 14) +
    labs(
        title = "Index Year",
        x     = "Year",
        y     = "Density",
        fill  = "Group",
        color = "Group"
    )
```

**KS test — index year:**

```{r ks-year}
print(r$ks_year)
```

---

# Data Integrity Check

```{r uniqueness-check}
if (r$uniqueness_check) {
    cat(sprintf("PASS: Each patient appears in exactly one selected %s period row.\n", params$nontreatment_group))
} else {
    cat(sprintf("WARNING: Some patients appear in more than one selected %s period row.\n", params$nontreatment_group))
}
```
